---
title: "Split-Apply-Combine..."
author: "Akshay Yadav"
topic: "02"
layout: post
root: ../../../
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
```

One of the best advantages of R language over other scripting languages like PERL or PYTHON are the abilities to iterate over multidimentional arrays or lists without the annoying FOR/WHILE loops. For example, if we want calculate column-wise means for some tabular data in PERL or PYTHON, a nested for loop for rows inside a for loop for columns is required. 
This can be easily achieved in R without the FOR/WHILE loops using the base functions, which split the matrix columns-wise; apply the mean function on each column and return the combined means of all the columns in form of an array. The R base functions that can use this split-apply-combine strategy on data structures are the functions from the apply family - lapply, sapply, mapply and other functions like aggregate
The main limitation of these is that they require specific input formats and the output formats are restricted. For example, lapply will only return a list. Also, the R base functions cannot be used easily, if splitting on multiple columns is required.

We will study the split-apply-combine strategy used by different functions on the **Glass** dataset from the **mlbench** package
```{r}
library(plyr)
library(dplyr)
data(Glass,package = "mlbench")
```

There are 7 different Types of Glass in the dataset with other 9 columns giving the amount of different elements present in the corresponding type

```{r}
Glass %>% glimpse()
table(Glass$Type)
```

We will calculate the mean of each element present in each type of Glass using the split-apply-combine strategy where first we will split data- rows-wise - according to the Glass type, apply the mean function to each portion of the data, and then combine the results

### Using the **aggregate** function
```{r}
aggregate(Glass[1:9],list(Type=Glass$Type),mean)
```

### Using the **ddply** from **ply** package
```{r}
ddply(Glass,.(Type),summarize,mean.RI=mean(RI),mean.Na=mean(Na),mean.Mg=mean(Mg),mean.Al=mean(Al),mean.Si=mean(Si),mean.K=mean(K),mean.Ca=mean(Ca),mean.Ba=mean(Ba),mean.Fe=mean(Fe))
```

### Using functions from dplyr packages
```{r}
Glass %>% group_by(Type) %>% summarise (mean.RI=mean(RI),mean.Na=mean(Na),mean.Mg=mean(Mg),mean.Al=mean(Al),mean.Si=mean(Si),mean.K=mean(K),mean.Ca=mean(Ca),mean.Ba=mean(Ba),mean.Fe=mean(Fe))
```





 