---
title: "The S3 class system"
author: "Akshay Yadav"
topic: "05"
layout: post
root: ../../../
---

####Explain how a user can define a new S3 class with objects and methods at a small example of your choice

Classes are the central to implementation of Object Oriented Programming. These are kind of blue-prints for holding and processing particular types of data.

We will look at S3 class type object-orientation in R. We will create a generic class for holding information on different types of vehicles and define methods for extracting this information.  

First, we will define some instances for the class **vehicle**. This is a little different from the classic object-oriented programming languages like JAVA where we first define the class and then instantiate it

```{r}
car<-list(wheels=4, steering=TRUE, capacity=4)
bike<-list(wheels=2, steering=FALSE, capacity=2)
trailer<-list(wheels=6, steering=TRUE, capacity=6)
```

Assigning the class attribute **vehicle** to all the instances
```{r}
class(car)<-append(class(car),"vehicle")
class(bike)<-append(class(bike),"vehicle")
class(trailer)<-append(class(trailer),"vehicle")

class(car)
class(bike)
class(trailer)
```

We will define a method for getting the **number of wheels** for instances of class **vehicle**

```{r}
getNwheels <- function(x)
{
     UseMethod("getNwheels",x)
}
 getNwheels.vehicle <- function(x)
 {
    return(x$wheels)
 }
```

Calling the **vehicle** class function *getNwheels()*
```{r}
getNwheels(car)
getNwheels(bike)
getNwheels(trailer)
```


